---
title: "dauerProtocol3_processing"
author: "Corinne Croslyn"
date: "5/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(EMCluster)
library(forcats)

# set working directory
setwd(glue::glue("{dirname(rstudioapi::getActiveDocumentContext()$path)}/.."))
```

## Part 1: Process Sorter Data

```{r}
# Define a vector of your experiment directories. Can load more
dirs <- c("data/raw/20220427_dauerProtocol3") 

# Read in the data using easysorter: https://github.com/AndersenLab/easysorter
raw <- easysorter::read_data(dirs)

# look at score data for each color channel
score_proc <- as.data.frame(raw[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain)) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = "")) #%>%

#renaming categorical variables in our condition column
score_proc$condition[score_proc$condition == "4hr_20uMascr"] <- "4hr 20 uM ascr#5"
score_proc$condition[score_proc$condition == "4hr_10uMascr"] <- "4hr 10 uM ascr#5"
score_proc$condition[score_proc$condition == "4hr_5uMascr"] <- "4hr 5 uM ascr#5"
score_proc$condition[score_proc$condition == "4hr_control"] <- "4hr control"

score_proc$condition[score_proc$condition == "20min_20uMascr"] <- "20min 20 uM ascr#5"
score_proc$condition[score_proc$condition == "20min_10uMascr"] <- "20min 10 uM ascr#5"
score_proc$condition[score_proc$condition == "20min_5uMascr"] <- "20min 5 uM ascr#5"
score_proc$condition[score_proc$condition == "20min_control"] <- "20min control"
```


## Part 2: Use Daehan's Data to get dauer classes (from green channel)

```{r}
raw_dauersGWA <- easysorter::read_data("data/raw/20170917_dauerGWA3A") # FOR NOW USE THIS ASSAY FOR SETTINGS

proc_dauersGWA <- as.data.frame(raw_dauersGWA[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain)) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

# EM clustering
df_raw <- proc_dauersGWA %>%
  dplyr::filter(!is.na(strain)) %>%
  dplyr::filter(condition == "ascr5.800nM") %>%
  dplyr::filter(TOF>100)

D_green <- df_raw[c("TOF", "green", "EXT")]
ret.em_K4 <- EMCluster::init.EM(D_green, nclass = 4, method = "em.EM")
D_green_merge_K4 <- data.frame(df_raw, ret.em_K4[6])

# plot to define classes
set.seed(5)
ggplot(D_green_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "GWA3a 800nM ascr#5", 
       subtitle = "Using Daehan's green channel data to define classes") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```

```{r}
#CAUTION!!! Labels change every time!!
D_green_merge_K4$class <- factor(D_green_merge_K4$class, levels=c(3,1,2,4), labels=c("dauer", "non-dauer", "other1", "other2"))

D_green_merge_K4$class <- fct_collapse(D_green_merge_K4$class, dauer = c("dauer"), nondauer = c("non-dauer"), other = c("other1", "other2"))

# replot with new factor names
gwa3_plot_green_channel <- D_green_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = class) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "GWA3a 800nM ascr#5",
       subtitle = "Using Daehan's green channel data to define classes",
       color = "") +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

gwa3_plot_green_channel

cowplot::ggsave2(gwa3_plot_green_channel, filename = "plots/dauerAssay3/GWA3a_classes_green_channel.png", width =5, height = 5)
```


## Part 3: Try to apply those (green) classes to our data

### Looking at green channel
```{r}
# applying model from DI condition to all our data
ours_g <- score_proc[c("TOF","green", "EXT")]
ret.em_K4_ours_g <- EMCluster::assign.class(ours_g, ret.em_K4)
y_merge_K4_ours_g <- data.frame(score_proc, ret.em_K4_ours_g[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_g$class <- factor(y_merge_K4_ours_g$class, levels=c(3,1,2,4), labels=c("dauer", "non-dauer", "other1", "other2"))

y_merge_K4_ours_g$class <- fct_collapse(y_merge_K4_ours_g$class, dauer = c("dauer"), nondauer = c("non-dauer"), other = c("other1", "other2"))

#ordering our categorical variables 
y_merge_K4_ours_g$condition <- factor(y_merge_K4_ours_g$condition, levels = c("20min control","4hr control", "20min 5 uM ascr#5", "4hr 5 uM ascr#5", "20min 10 uM ascr#5", "4hr 10 uM ascr#5", "20min 20 uM ascr#5", "4hr 20 uM ascr#5"), ordered=TRUE)

# plot to define classes
dauer_assay3_green_daehan <- ggplot(y_merge_K4_ours_g) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay #3 (Green Channel)", 
       subtitle = "Applying Daehan's classes to our data",
       color = "") +
  facet_wrap(~condition, ncol = 2) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_green_daehan

cowplot::ggsave2(dauer_assay3_green_daehan, filename = "plots/dauerAssay3/dauer_assay3_daehan_clusters_by_condition_green.png", width = 5, height = 10)
```

Daehan's green channel classes doesn't seem to be very effective at predicting our dauer classes, because we see a bimodal non-dauer pattern.

## Part 4: Use Daehan's Data to get dauer classes (from red channel)

```{r}
D_red <- df_raw[c("TOF", "red", "EXT")]
#using just two classes here as it tends to split the dauer class in half at the bottom with 3-4 classes
ret.em_K4_2 <- EMCluster::init.EM(D_red, nclass = 4, method = "em.EM")
D_red_merge_K4 <- data.frame(df_raw, ret.em_K4_2[6])

# plot to define classes
ggplot(D_red_merge_K4) +
  aes(x = TOF, y = red, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "GWA3a 800nM ascr#5") +
  facet_wrap(~condition, ncol = 2) +
  theme_bw() +
  ylim(0,300) +
  xlim(0,600)
```


```{r}
D_red_merge_K4$class <- factor(D_red_merge_K4$class, levels=c(4,2,1,3), labels=c("dauer", "non-dauer", "other1", "other2"))

D_red_merge_K4$class <- fct_collapse(D_red_merge_K4$class, dauer = c("dauer"), nondauer = c("non-dauer"), other = c("other1", "other2"))

# replot with new factor names
gwa3_plot_red <- D_red_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = red, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "GWA3a 800nM ascr#5", color = "") +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

gwa3_plot_red

cowplot::ggsave2(gwa3_plot_red, filename = "plots/dauerAssay3/GWA3a_red_classes.png", width =5, height = 5)
```

## Part 5: Try to apply Daehan's (red) classes to our data 

### Looking at our red channel
```{r}
# applying model from DI condition to all our data
ours_r <- score_proc[c("TOF","red", "EXT")]
ret.em_K4_ours_r <- EMCluster::assign.class(ours_r, ret.em_K4_2)
y_merge_K4_ours_r <- data.frame(score_proc, ret.em_K4_ours_r[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_r$class <- factor(y_merge_K4_ours_r$class, levels=c(4,2,1,3), labels=c("dauer", "non-dauer", "other1", "other2"))

y_merge_K4_ours_r$class <- fct_collapse(y_merge_K4_ours_r$class, dauer = c("dauer"), nondauer = c("non-dauer"), other = c("other1", "other2"))

#ordering our categorical variables 
y_merge_K4_ours_r$condition <- factor(y_merge_K4_ours_r$condition, levels = c("20min control","4hr control", "20min 5 uM ascr#5", "4hr 5 uM ascr#5", "20min 10 uM ascr#5", "4hr 10 uM ascr#5", "20min 20 uM ascr#5", "4hr 20 uM ascr#5"), ordered=TRUE)

# plot to define classes
dauer_assay3_red <- ggplot(y_merge_K4_ours_r) +
  aes(x = TOF, y = red, scale = T, color = as.factor(class)) +
  #scale_color_manual(values = ) # can use to set colors consistently
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay #3 (Red Channel)", 
       subtitle = "Applying Daehan's classes to our data",
       color = "") +
  facet_wrap(~condition, ncol = 2) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_red

cowplot::ggsave2(dauer_assay3_red, filename = "plots/dauerAssay3/dauer_assay3_daehan_clusters_by_condition_red.png", width = 5, height = 10)
```

Daehan's red channel classes also don't predict our dauers very well. Our clusters look completely different.  
The red data seems to be all over the place in our data, while Daehan's is much cleaner.  

We should try training our data using a subset of Dauer Assay #3 and then applying those classes to the whole data. 
## Part 6: Using our controls to train clustering of our (green) data 

### Getting classes from our own (green) data

```{r}
raw_dauers_ours <- easysorter::read_data("data/raw/20220427_dauerProtocol3") 

proc_dauers_ours <- as.data.frame(raw_dauers_ours[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain)) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

#renaming categorical variables in our condition column
proc_dauers_ours$condition[proc_dauers_ours$condition == "4hr_20uMascr"] <- "4hr 20 uM ascr#5"
proc_dauers_ours$condition[proc_dauers_ours$condition == "4hr_10uMascr"] <- "4hr 10 uM ascr#5"
proc_dauers_ours$condition[proc_dauers_ours$condition == "4hr_5uMascr"] <- "4hr 5 uM ascr#5"
proc_dauers_ours$condition[proc_dauers_ours$condition == "4hr_control"] <- "4hr control"

proc_dauers_ours$condition[proc_dauers_ours$condition == "20min_20uMascr"] <- "20min 20 uM ascr#5"
proc_dauers_ours$condition[proc_dauers_ours$condition == "20min_10uMascr"] <- "20min 10 uM ascr#5"
proc_dauers_ours$condition[proc_dauers_ours$condition == "20min_5uMascr"] <- "20min 5 uM ascr#5"
proc_dauers_ours$condition[proc_dauers_ours$condition == "20min_control"] <- "20min control"

# EM clustering in DI condition
df_raw_ours <- proc_dauers_ours %>%
  dplyr::filter(!is.na(strain)) %>%
  dplyr::filter(condition %in% c("4hr control", "20min control", "4hr 20 uM ascr#5", "20min 20 uM ascr#5")) %>%
  dplyr::filter(TOF>100)

ours_green <- df_raw_ours[c("TOF", "green", "EXT")]
ret.em_K4_ours_g2 <- EMCluster::init.EM(ours_green, nclass = 3, method = "em.EM")
ours_green2_merge_K4 <- data.frame(df_raw_ours, ret.em_K4_ours_g2[6])

# plot to define classes
ggplot(ours_green2_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay #3 (Green Channel)") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```

```{r}
# Caution!!! factor label changes everytime!!!
ours_green2_merge_K4$class <- factor(ours_green2_merge_K4$class, levels=c(1,2,3), labels=c("dauer","non-dauer","other1"))

ours_green2_merge_K4$class <- fct_collapse(ours_green2_merge_K4$class, dauer = c("dauer"), nondauer = c("non-dauer", "other1"))

# replot with new factor names
ours_plot_green <- ours_green2_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "Dauer Assay #3 (Green Channel)",
       subtitle = "Using subset to define classes",
       color = "") +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

ours_plot_green

cowplot::ggsave2(ours_plot_green, filename = "plots/dauerAssay3/our_green_classes_green.png", width =5, height = 5)
```

### Using our (green) classes (defined from our own data) to apply to the rest of our (green) data

```{r}
# applying model from DI condition to all our data
ours_g2 <- proc_dauers_ours[c("TOF","green", "EXT")]
ret.em_K4_ours_g2.1 <- EMCluster::assign.class(ours_g2, ret.em_K4_ours_g2) 
y_merge_K4_ours_g2.1 <- data.frame(proc_dauers_ours, ret.em_K4_ours_g2.1[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_g2.1$class <- factor(y_merge_K4_ours_g2.1$class, levels=c(1,2,3), labels=c("dauer","non-dauer","other1"))

y_merge_K4_ours_g2.1$class <- fct_collapse(y_merge_K4_ours_g2.1$class, dauer = c("dauer"), nondauer = c("non-dauer", "other1"))

#ordering our categorical variables 
y_merge_K4_ours_g2.1$condition <- factor(y_merge_K4_ours_g2.1$condition, levels = c("20min control","4hr control", "20min 5 uM ascr#5", "4hr 5 uM ascr#5", "20min 10 uM ascr#5", "4hr 10 uM ascr#5", "20min 20 uM ascr#5", "4hr 20 uM ascr#5"), ordered=TRUE)

# plot to define classes
dauer_assay3_green <- ggplot(y_merge_K4_ours_g2.1) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 3 Classes (Green Channel)", 
       subtitle = "Using a subset to define classes",
       color = "") +
  facet_wrap(~condition, ncol=2) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_green

cowplot::ggsave2(dauer_assay3_green, filename = "plots/dauerAssay3/dauer_assay3_our_clusters_by_condition_green.png", width = 5, height = 10)
```
## Part 7: Using our controls to train clustering of our own (red) data

### Getting classes from our red channel

```{r}
ours_red <- df_raw_ours[c("TOF", "red", "EXT")]
ret.em_K4_ours_r2 <- EMCluster::init.EM(ours_red, nclass = 3, method = "em.EM")
ours_red2_merge_K4 <- data.frame(df_raw_ours, ret.em_K4_ours_r2[6])

# plot to define classes
ggplot(ours_red2_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay #3 (Red Channel)") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```
```{r}
# Caution!!! factor label changes everytime!!!
ours_red2_merge_K4$class <- factor(ours_red2_merge_K4$class, levels=c(3,2,1), labels=c("dauer","non-dauer","other1"))

ours_red2_merge_K4$class <- fct_collapse(ours_red2_merge_K4$class, dauer = c("dauer"), nondauer = c("non-dauer"), other =c("other1"))

# replot with new factor names
ours_plot_red <-ours_red2_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "Dauer Assay #3 (Red Channel)",
       subtitle = "Using subset to define classes",
       color = "") +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

ours_plot_red

cowplot::ggsave2(ours_plot_red, filename = "plots/dauerAssay3/our_red_classes_.png", width =5, height = 5)
```

### Using our (red) classes (defined from our own data) to apply to the rest of our (red) data

```{r}
# applying model from DI condition to all our data
ours_r2 <- proc_dauers_ours[c("TOF","red", "EXT")]
ret.em_K4_ours_r2.1 <- EMCluster::assign.class(ours_r2, ret.em_K4_ours_r2) 
y_merge_K4_ours_r2.1 <- data.frame(proc_dauers_ours, ret.em_K4_ours_r2.1[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_r2.1$class <- factor(y_merge_K4_ours_r2.1$class, levels=c(3,2,1), labels=c("dauer","non-dauer","other1"))

y_merge_K4_ours_r2.1$class <- fct_collapse(y_merge_K4_ours_r2.1$class, dauer = c("dauer"), nondauer = c("non-dauer", "other1"))

#ordering our categorical variables 
y_merge_K4_ours_g2.1$condition <- factor(y_merge_K4_ours_g2.1$condition, levels = c("20min control","4hr control", "20min 5 uM ascr#5", "4hr 5 uM ascr#5", "20min 10 uM ascr#5", "4hr 10 uM ascr#5", "20min 20 uM ascr#5", "4hr 20 uM ascr#5"), ordered=TRUE)

# plot to define classes
dauer_assay3_red <- ggplot(y_merge_K4_ours_r2.1) +
  aes(x = TOF, y = red, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay #3 Classes (Red Channel)", 
       subtitle = "Using a subset to define classes",
       color = "") +
  facet_wrap(~condition, ncol=2) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_red

cowplot::ggsave2(dauer_assay3_red, filename = "plots/dauerAssay3/dauer_assay3_our_clusters_by_condition_red.png", width = 5, height = 10)
```

#===============================================================================#
# make box plots
#===============================================================================#
da1_sum <- y_merge_K4_ours1 %>%
  dplyr::group_by(well) %>%
  dplyr::mutate(well_tot_n = n(),
                well_dauer_n = sum(class == "dauer"),
                well_nondauer_n = sum(class == "non-dauer"),
                well_other1_n = sum(class == "other1"),
                dauerFrac1 = well_dauer_n/(well_nondauer_n + well_dauer_n),
                dauerFrac2 = well_dauer_n/(well_other1_n + well_dauer_n))

da2_sum <- y_merge_K4_ours %>%
  dplyr::group_by(well) %>%
  dplyr::mutate(well_tot_n = n(),
                well_dauer_n = sum(class == "dauer"),
                well_nondauer_n = sum(class == "non-dauer"),
                well_other1_n = sum(class == "other1"),
                dauerFrac1 = well_dauer_n/(well_nondauer_n + well_dauer_n),
                dauerFrac2 = well_dauer_n/(well_other1_n + well_dauer_n))

# box plots
bp1 <- ggplot(da1_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac2, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 1) +
  labs(y = "dauer fraction (d/(d+other1))", title = "Assay1") +
  theme(legend.position = "none")
bp1

bp1b <- ggplot(da1_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac1, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 1) +
  labs(y = "dauer fraction (d/(d+nond))", title = "Assay1") +
  theme(legend.position = "none")
bp1b

bp2a <- ggplot(da2_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac1, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 2) +
  labs(y = "dauer fraction (d/(d+nond))", title = "Assay2") +
  theme(legend.position = "none")
bp2a

bp2b <- ggplot(da2_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac2, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 2) +
  labs(y = "dauer fraction (d/(d+other1))", title = "Assay2") +
  theme(legend.position = "none")
bp2b

cowplot::ggsave2(bp1, filename = "plots/assay1_dauerF_boxplot1.png", width = 5, height = 5)
cowplot::ggsave2(bp1b, filename = "plots/assay1_dauerF_boxplot2.png", width = 5, height = 5)
cowplot::ggsave2(bp2a, filename = "plots/assay2_dauerF_boxplot1.png", width = 5, height = 5)
cowplot::ggsave2(bp2b, filename = "plots/assay2_dauerF_boxplot2.png", width = 5, height = 5)
