---
title: "dauerProtocol5_processing"
author: "Corinne Croslyn"
date: '2022-05-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(EMCluster)
library(forcats)

# set working directory
setwd(glue::glue("{dirname(rstudioapi::getActiveDocumentContext()$path)}/.."))
```


## Part 1: process sorter data

```{r}
# Define a vector of your experiment directories. Can load more
dirs <- c("data/raw/20220518_dauerAssay5") 

# Read in the data using easysorter: https://github.com/AndersenLab/easysorter
raw <- easysorter::read_data(dirs)

# look at score data for each color channel
score_proc <- as.data.frame(raw[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain), TOF>100) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

#renaming categorical variables in our condition column
score_proc$condition[score_proc$condition == "4hr_20uMascr"] <- "4hr 20 uM ascr#5"
score_proc$condition[score_proc$condition == "4hr_10uMascr"] <- "4hr 10 uM ascr#5"
score_proc$condition[score_proc$condition == "4hr_5uMascr"] <- "4hr 5 uM ascr#5"
score_proc$condition[score_proc$condition == "4hr_control"] <- "4hr control"

score_proc$condition[score_proc$condition == "20min_20uMascr"] <- "20min 20 uM ascr#5"
score_proc$condition[score_proc$condition == "20min_10uMascr"] <- "20min 10 uM ascr#5"
score_proc$condition[score_proc$condition == "20min_5uMascr"] <- "20min 5 uM ascr#5"
score_proc$condition[score_proc$condition == "20min_control"] <- "20min control"
```

## Part 2: Using our controls to define clustering of our data

### Getting classes from our own (green) data

```{r}
# EM clustering
df_raw <- score_proc %>%
  dplyr::filter(!is.na(strain)) %>%
  dplyr::filter(condition %in% c("4hr control","4hr 20 uM ascr#5")) %>%
  dplyr::filter(TOF>100)

df_green <- df_raw[c("TOF", "green", "EXT")]
ret.em_K4_green <- EMCluster::init.EM(df_green, nclass = 5, method = "em.EM") 
green_merge_K4 <- data.frame(df_raw, ret.em_K4_green[6])

# plot to define classes
set.seed(5)
raw_classes2 <- ggplot(green_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay #3 Classifiers",
       subtitle = "Using 4hr control + 20min control + 4hr 20 uM + 20min 20 uM") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)

cowplot::ggsave2(raw_classes2, filename = "plots/dauerAssay3/raw_green_classes1.png", width =5, height = 5)
```

```{r}
# Caution!!! factor label changes every time!!!
green_merge_K4$class <- factor(green_merge_K4$class, levels=c(1,5,4,2,3), labels=c("dauer", "non-dauer1","non-dauer2", "other1", "other2"))

green_merge_K4$class <- fct_collapse(green_merge_K4$class, dauer = c("dauer", "other1"), nondauer = c("non-dauer1", "non-dauer2"), other = c("other2"))

# replot with new factor names
ours_plot_green <- green_merge_K4 %>%
  #dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "Dauer Assay 3 20uM ascr5 + 10 uM ascr5 + Control, 4hr (green)", color = "") +
  theme_bw() +
 #ylim(0, 300) +
  xlim(0, 600)

ours_plot_green

cowplot::ggsave2(ours_plot_green, filename = "plots/dauerAssay3/our_green_classes.png", width =5, height = 5)
```

### Part 3: Using our classes (defined from our own data) to apply to the rest of our data

```{r, fig.height=15}
# applying model from DI condition to all our data
ours_g <- score_proc[c("TOF","green", "EXT")] %>%
  dplyr::filter(TOF>100)
ret.em_K4_ours_g <- EMCluster::assign.class(ours_g, ret.em_K4_green) 
y_merge_K4_ours_g <- data.frame(score_proc, ret.em_K4_ours_g[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_g$class <- factor(y_merge_K4_ours_g$class, levels=c(1,5,4,2,3), labels=c("dauer", "non-dauer1","non-dauer2", "other1", "other2"))

y_merge_K4_ours_g$class <- fct_collapse(y_merge_K4_ours_g$class, dauer = c("dauer", "other1"), nondauer = c("non-dauer1", "non-dauer2"), other = c("other2"))

#ordering our categorical variables 
y_merge_K4_ours_g$condition <- factor(y_merge_K4_ours_g$condition, levels = c("20min control","4hr control", "20min 5 uM ascr#5", "4hr 5 uM ascr#5", "20min 10 uM ascr#5", "4hr 10 uM ascr#5", "20min 20 uM ascr#5", "4hr 20 uM ascr#5"), ordered=TRUE)

# plot to define classes
dauer_assay3_green <- ggplot(y_merge_K4_ours_g) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  #scale_color_manual(values = ) # can use to set colors consistently
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 3 Classes (Green Channel)", color = "") +
  facet_wrap(~condition, ncol = 2) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_green

cowplot::ggsave2(dauer_assay3_green, filename = "plots/dauerAssay3/dauer_assay3_clusters_by_condition_green.png", width = 10, height = 10)
```

## Part 4: Quantifying dauers

### Calculating the dauer fraction

```{r}
#calculating the dauer fraction in each well
da1_sum <- y_merge_K4_ours_g %>%
  dplyr::group_by(well) %>%
  dplyr::mutate(well_tot_n = n(),
                well_dauer_n = sum(class == "dauer"),
                well_nondauer_n = sum(class == "nondauer"),
                well_other_n = sum(class == "other"),
                dauerFrac1 = well_dauer_n/(well_nondauer_n + well_dauer_n),
                dauerFrac2 = well_dauer_n/(well_other_n + well_dauer_n))

#calculating the average amount of worms in each well (should be ~50)
mean(da1_sum$well_tot_n)

#calculating the dauerfrac means for each condition
dauerfrac_mean <- da1_sum %>%
  dplyr::group_by(condition) %>%
  dplyr::mutate(condition_mean = sum(class == "dauer")/n()) %>%
  dplyr::distinct(condition, condition_mean)
```
The condition means clearly show a dose-dependent response!

The average worm/well in this assay was ~32. This is low, but I think
our pipetting improved after this. The high was 47, the low was 18; this
is probably just an inconsistency in pipetting problem.

### Graphing

```{r, fig.height=15}
# box plots
bp1 <- ggplot(da1_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac2, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 2) +
  labs(y = "dauer fraction (d/(d+other))", title = "Dauer Assay 3") +
  theme(legend.position = "none")
bp1

bp1b <- ggplot(da1_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac1, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  geom_text(data = aggregate(dauerFrac1~well, da1_sum, mean)) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 2) +
  labs(y = "dauer fraction (d/(d+nond))", title = "Dauer Assay 3") +
  theme(legend.position = "none")
bp1b

cowplot::ggsave2(bp1, filename = "plots/dauerAssay3/assay3_dauerF_boxplot1.png", width = 5, height = 10)
cowplot::ggsave2(bp1b, filename = "plots/dauerAssay3/assay3_dauerF_boxplot2.png", width = 5, height = 10)
```
