---
title: "dauerProtocol3_processing"
author: "Corinne Croslyn"
date: "5/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(EMCluster)
library(forcats)

# set working directory
setwd(glue::glue("{dirname(rstudioapi::getActiveDocumentContext()$path)}/.."))
```

## Part 1: Process Sorter Data

```{r}
# Define a vector of your experiment directories. Can load more
dirs <- c("data/raw/20220427_dauerProtocol3") 

# Read in the data using easysorter: https://github.com/AndersenLab/easysorter
raw <- easysorter::read_data(dirs)

# look at score data for each color channel
score_proc <- as.data.frame(raw[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain)) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = "")) #%>%
#tidyr::pivot_longer(cols = c(green, yellow, red))

# # plot Fl by TOF for all channels
# a1_channels <- ggplot(score_proc %>% dplyr::filter(!is.na(condition) & !is.na(strain))) +
#   aes(x = TOF, y = value, fill = condition) +
#   geom_point(size=1, alpha = 0.3, shape = 21) +
#   #facet_wrap(~condition) +
#   labs(y = "Fluorescence (AU)") +
#   facet_grid(strain~condition~name) +
#   theme_bw() +
#   theme(legend.position = "none")
# a1_channels

#cowplot::ggsave2(a1_channels, filename = "plots/dauerProtocol1_sorter_FLchannels.png")
```


## Part 2: Use Daehan's Data to get dauer classes (from green channel)

```{r}
raw_dauersGWA <- easysorter::read_data("data/raw/20170917_dauerGWA3A") # FOR NOW USE THIS ASSAY FOR SETTINGS

proc_dauersGWA <- as.data.frame(raw_dauersGWA[1]) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

# EM clustering in DI condition
df_raw <- proc_dauersGWA %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(strain %in% c("JU346", "ECA1089", "ECA1119", "ECA1121", "NIC166", "ECA1091", "ECA1120", "ECA1122")) %>%
  dplyr::filter(condition == "ascr5.800nM") %>%
  dplyr::filter(TOF>100)

D_green <- df_raw[c("TOF", "green", "EXT")]
ret.em_K4 <- EMCluster::init.EM(D_green, nclass = 4, method = "em.EM")
D_green_merge_K4 <- data.frame(df_raw, ret.em_K4[6])

# plot to define classes
ggplot(D_green_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "GWA3a 800nM ascr#5") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```

```{r}
# Caution!!! factor label changes everytime!!!
# labels=c("crap1", "crap2", "non-dauer", "dauer")
# labels=c("crap1", "dauer",  "crap2", "non-dauer")
# labels=c("non-dauer", "crap1", "crap2", "dauer")
D_green_merge_K4$class <- factor(D_green_merge_K4$class, levels=c(1,4,2,3), labels=c("dauer", "non-dauer", "other1", "other2"))

# replot with new factor names
gwa3_plot_green_channel <- D_green_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = class) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "GWA3a 800nM ascr#5", color = "") +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

gwa3_plot_green_channel

cowplot::ggsave2(gwa3_plot_green_channel, filename = "plots/GWA3a_classes_green_channel.png", width =5, height = 5)
```


## Part 3: Try to apply those (green) classes to our data

### Looking at green channel
```{r}
# applying model from DI condition to all our data
ours_g <- score_proc[c("TOF","green", "EXT")]
ret.em_K4_ours_g <- EMCluster::assign.class(ours_g, ret.em_K4)
y_merge_K4_ours_g <- data.frame(score_proc, ret.em_K4_ours_g[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_g$class <- factor(y_merge_K4_ours_g$class, levels=c(1,4,2,3), labels=c("dauer", "non-dauer", "other1", "other2"))

# plot to define classes
dauer_assay3_green <- ggplot(y_merge_K4_ours_g) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  #scale_color_manual(values = ) # can use to set colors consistently
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 3 Classes (Green Channel)", color = "") +
  facet_wrap(~condition) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_green

cowplot::ggsave2(dauer_assay3_green, filename = "plots/dauer_assay3_clusters_by_condition_green.png", width = 5, height = 5)
```
Daehan's green channel classes doesn't seem to be very effective at predicting our dauer classes

## Part 4: Use Daehan's Data to get dauer classes (from red channel)

```{r}
D_red <- df_raw[c("TOF", "red", "EXT")]
#using just two classes here as it tends to split the dauer class in half at the bottom with 3-4 classes
ret.em_K4_2 <- EMCluster::init.EM(D_red, nclass = 3, method = "em.EM")
D_red_merge_K4 <- data.frame(df_raw, ret.em_K4_2[6])

# plot to define classes
ggplot(D_red_merge_K4) +
  aes(x = TOF, y = red, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "GWA3a 800nM ascr#5") +
  theme_bw() 
```
```{r}
# Caution!!! factor label changes everytime!!!
# labels=c("crap1", "crap2", "non-dauer", "dauer")
# labels=c("crap1", "dauer",  "crap2", "non-dauer")
# labels=c("non-dauer", "crap1", "crap2", "dauer")
D_red_merge_K4$class <- factor(D_red_merge_K4$class, levels=c(2,3,1), labels=c("dauer", "non-dauer", "other"))

# replot with new factor names
gwa3_plot_red <- D_red_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = red, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "GWA3a 800nM ascr#5", color = "") +
  theme_bw() +
 #ylim(0, 300) +
  xlim(0, 600)

gwa3_plot_red

cowplot::ggsave2(gwa3_plot_red, filename = "plots/GWA3a_red_classes.png", width =5, height = 5)
```

## Part 5: Try to apply Daehan's (red) classes to our data 

### Looking at our red channel
```{r}
# applying model from DI condition to all our data
ours_r <- score_proc[c("TOF","red", "EXT")]
ret.em_K4_ours_r <- EMCluster::assign.class(ours_r, ret.em_K4)
y_merge_K4_ours_r <- data.frame(score_proc, ret.em_K4_ours_r[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_r$class <- factor(y_merge_K4_ours_r$class, levels=c(2,3,1), labels=c("dauer", "non-dauer", "other"))

# plot to define classes
dauer_assay3_red <- ggplot(y_merge_K4_ours_r) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  #scale_color_manual(values = ) # can use to set colors consistently
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 3 Classes (Red Channel)", color = "") +
  facet_wrap(~condition) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_red

cowplot::ggsave2(dauer_assay3_red, filename = "plots/dauer_assay3_clusters_by_condition_red.png", width = 5, height = 5)
```

Daehan's red channel classes also don't predict our dauers very well. Our clusters look completely different.  
The green and red clustering in our data does look almost identical though. 

We should try training our data by taking our control (no ascr) condition and using it to establish non-dauer, and then using that to cluster the rest of the data. 

## Part 6: Using our controls to train clustering of our data 

### Getting classes from our own (green) data

```{r}
raw_dauers_ours <- easysorter::read_data("data/raw/20220427_dauerProtocol3") # FOR NOW USE THIS ASSAY FOR SETTINGS

proc_dauers_ours <- as.data.frame(raw_dauers_ours[1]) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

# EM clustering in DI condition
df_raw_ours <- proc_dauers_ours %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(strain %in% c("JU346", "ECA1089", "ECA1119", "ECA1121", "NIC166", "ECA1091", "ECA1120", "ECA1122")) %>%
  dplyr::filter(condition %in% c("control_4hr", "20uM_ascr5_4hr", "10uM_ascr5_4hr"))
  #dplyr::filter(TOF>100)

ours_green <- df_raw_ours[c("TOF", "green", "EXT")]
ret.em_K4_ours_g2 <- EMCluster::init.EM(ours_green, nclass = 6, method = "em.EM")
ours_green2_merge_K4 <- data.frame(df_raw_ours, ret.em_K4_ours_g2[6])

# plot to define classes
ggplot(ours_green2_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 3 20uM ascr5 + Control, 4hr (green)") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```

```{r}
# Caution!!! factor label changes everytime!!!
# labels=c("crap1", "crap2", "non-dauer", "dauer")
# labels=c("crap1", "dauer",  "crap2", "non-dauer")
# labels=c("non-dauer", "crap1", "crap2", "dauer")
ours_green2_merge_K4$class <- factor(ours_green2_merge_K4$class, levels=c(1,4,), labels=c("dauer1","dauer2", "non-dauer1", "non-dauer2", "other1", "other2"))

ours_green2_merge_K4$class <- fct_collapse(ours_green2_merge_K4$class, dauer = c("dauer1","dauer2"), nondauer = c("non-dauer1", "non-dauer2"), other = c("other1","other2"))

# replot with new factor names
ours_plot_green <- ours_green2_merge_K4 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "GWA3a 800nM ascr#5", color = "") +
  theme_bw() +
 #ylim(0, 300) +
  xlim(0, 600)

ours_plot_green

cowplot::ggsave2(ours_plot_green, filename = "plots/our_green_classes.png", width =5, height = 5)
```

### Using our classes (defined from our own data) to apply to the rest of our data

```{r}
# applying model from DI condition to all our data
ours_g2 <- score_proc[c("TOF","green", "EXT")]
ret.em_K4_ours_g2.1 <- EMCluster::assign.class(ours_g2, ret.em_K4_ours_g2) 
y_merge_K4_ours_g2.1 <- data.frame(score_proc, ret.em_K4_ours_g2.1[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_ours_g2.1$class <- factor(y_merge_K4_ours_g2$class, levels=c(1,3,2,5,4,6), labels=c("dauer1","dauer2", "non-dauer1", "non-dauer2", "other1", "other2"))

y_merge_K4_ours_g2.1$class <- fct_collapse(y_merge_K4_ours_g2$class, dauer = c("dauer1","dauer2"), nondauer = c("non-dauer1", "non-dauer2"), other = c("other1","other2"))

# plot to define classes
dauer_assay3_green <- ggplot(y_merge_K4_ours_g2.1) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  #scale_color_manual(values = ) # can use to set colors consistently
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 3 Classes (Green Channel)", color = "") +
  facet_wrap(~condition) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

dauer_assay3_green

cowplot::ggsave2(dauer_assay3_red, filename = "plots/dauer_assay3_clusters_by_condition_green.png", width = 5, height = 5)
```

#===============================================================================#
# make box plots
#===============================================================================#
da1_sum <- y_merge_K4_ours1 %>%
  dplyr::group_by(well) %>%
  dplyr::mutate(well_tot_n = n(),
                well_dauer_n = sum(class == "dauer"),
                well_nondauer_n = sum(class == "non-dauer"),
                well_other1_n = sum(class == "other1"),
                dauerFrac1 = well_dauer_n/(well_nondauer_n + well_dauer_n),
                dauerFrac2 = well_dauer_n/(well_other1_n + well_dauer_n))

da2_sum <- y_merge_K4_ours %>%
  dplyr::group_by(well) %>%
  dplyr::mutate(well_tot_n = n(),
                well_dauer_n = sum(class == "dauer"),
                well_nondauer_n = sum(class == "non-dauer"),
                well_other1_n = sum(class == "other1"),
                dauerFrac1 = well_dauer_n/(well_nondauer_n + well_dauer_n),
                dauerFrac2 = well_dauer_n/(well_other1_n + well_dauer_n))

# box plots
bp1 <- ggplot(da1_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac2, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 1) +
  labs(y = "dauer fraction (d/(d+other1))", title = "Assay1") +
  theme(legend.position = "none")
bp1

bp1b <- ggplot(da1_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac1, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 1) +
  labs(y = "dauer fraction (d/(d+nond))", title = "Assay1") +
  theme(legend.position = "none")
bp1b

bp2a <- ggplot(da2_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac1, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 2) +
  labs(y = "dauer fraction (d/(d+nond))", title = "Assay2") +
  theme(legend.position = "none")
bp2a

bp2b <- ggplot(da2_sum %>% dplyr::distinct(well, .keep_all = T)) +
  aes(x = strain, 
      y = dauerFrac2, fill=strain,
      labs=strain) +
  ylim(0, 1)+
  geom_point(aes(fill=condition), position=position_jitterdodge(), alpha = .7) +
  geom_boxplot(outlier.alpha = 0, alpha = 0.7) +
  theme_bw() + 
  facet_wrap(~condition, ncol = 2) +
  labs(y = "dauer fraction (d/(d+other1))", title = "Assay2") +
  theme(legend.position = "none")
bp2b

cowplot::ggsave2(bp1, filename = "plots/assay1_dauerF_boxplot1.png", width = 5, height = 5)
cowplot::ggsave2(bp1b, filename = "plots/assay1_dauerF_boxplot2.png", width = 5, height = 5)
cowplot::ggsave2(bp2a, filename = "plots/assay2_dauerF_boxplot1.png", width = 5, height = 5)
cowplot::ggsave2(bp2b, filename = "plots/assay2_dauerF_boxplot2.png", width = 5, height = 5)
