---
title: "dauerProtocol4_processing"
author: "Corinne Croslyn"
date: "5/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(EMCluster)
library(forcats)

# set working directory
setwd(glue::glue("{dirname(rstudioapi::getActiveDocumentContext()$path)}/.."))
```

For dauer assay 4, we set up 2 plates (1 N2, 1 JU310) with our ascr dilution series (20, 10, 5, 0) and performed the assay the normal way (setting up the plate, feeding the same day, sorting after 48 hours). This assay was sorted on 20220504.

We also set up the same plates (1 N2, 1 JU3120 with ascr dilution series 20, 10, 5, 0), but didn't feed right away. Instead, we waited 24 hours to feed, then added food and allowed the worms to feed for 48 hr before sorting. This assay was sorted on 20220505. 

We want to see if there's a difference between feeding right away and waiting 24 hours

## Part 1: process sorter data from standard assay

```{r}
# Define a vector of your experiment directories. Can load more
dirs1 <- c("data/raw/20220504_dauerProtocol4") 

# Read in the data using easysorter: https://github.com/AndersenLab/easysorter
standard_raw <- easysorter::read_data(dirs1)

# look at score data for each color channel
score_proc1 <- as.data.frame(standard_raw[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain)) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))
```

## Part 2: Using our controls to train clustering of our standard data 

### Getting classes from our own (green) data

```{r}
standard_raw_dauers <- easysorter::read_data("data/raw/20220504_dauerProtocol4") # FOR NOW USE THIS ASSAY FOR SETTINGS

proc_dauers1 <- as.data.frame(standard_raw_dauers[1]) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

# EM clustering in DI condition
df_standard_raw <- proc_dauers1 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(strain %in% c("JU346", "ECA1089", "ECA1119", "ECA1121", "NIC166", "ECA1091", "ECA1120", "ECA1122")) %>%
  dplyr::filter(condition %in% c("20uM_ascr5"))
  #dplyr::filter(TOF>100)

standard_df_green <- df_standard_raw[c("TOF", "green", "EXT")]
ret.em_K4_green_standard <- EMCluster::init.EM(standard_df_green, nclass = 4, method = "em.EM")
standard_green_merge_K4 <- data.frame(df_standard_raw, ret.em_K4_green_standard[6])

# plot to define classes
ggplot(standard_green_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Dauer Assay 4 Class Training (green)") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```

```{r}
# Caution!!! factor label changes everytime!!!
standard_green_merge_K4$class <- factor(standard_green_merge_K4$class, levels=c(), labels=c("dauer","non-dauer", "other1", "other2"))

standard_green_merge_K4$class <- fct_collapse(standard_green_merge_K4$class, dauer = c("dauer"), nondauer = c("non-dauer1"), other = c("other1","other2"))

# replot with new factor names
standard_plot_green <- standard_green_merge_K4 %>%
  #dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "Standard Dauer Assay 4 Classes (green)", color = "") +
  theme_bw() +
 #ylim(0, 300) +
  xlim(0, 600)

standard_plot_green

cowplot::ggsave2(standard_plot_green, filename = "plots/standard_green_classes4.png", width =5, height = 5)
```

### Using our classes (defined from our own data) to apply to the rest of our data

```{r}
# applying model from DI condition to all our data
standard_g <- score_proc1[c("TOF","green", "EXT")]
ret.em_K4_standard_g <- EMCluster::assign.class(standard_g, ret.em_K4_green_standard) 
y_merge_K4_standard_g <- data.frame(score_proc1, ret.em_K4_standard_g[6])

# CHANGES EVERY TIME! (same as defined above)
y_merge_K4_standard_g$class <- factor(y_merge_K4_standard_g$class, levels=c(1,4,2,3,5,6), labels=c("dauer","non-dauer","other1", "other2"))

y_merge_K4_standard_g$class <- fct_collapse(y_merge_K4_ours_g2$class, dauer = c("dauer"), nondauer = c("non-dauer"), other = c("other1","other2"))

# plot to define classes
standard_dauer_assay4_green <- ggplot(y_merge_K4_standard_g) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  #scale_color_manual(values = ) # can use to set colors consistently
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "Standard Dauer Assay 4 Classes (Green Channel)", color = "") +
  facet_wrap(~condition) +
  theme_bw() +
  ylim(0, 300) +
  xlim(0, 600)

standard_dauer_assay4_green

cowplot::ggsave2(dauer_assay3_green, filename = "plots/standard_dauer_assay4_clusters_by_condition_green.png", width = 5, height = 5)
```

## Part 3: process sorter data from 24hr assay

```{r}
# Define a vector of your experiment directories. Can load more
dirs2 <- c("data/raw/20220505_dauerProtocol4") 

# Read in the data using easysorter: https://github.com/AndersenLab/easysorter
raw_24hr <- easysorter::read_data(dirs2)

# look at score data for each color channel
score_proc2 <- as.data.frame(raw_24hr[1]) %>%
  dplyr::filter(call50 != "bubble", contamination != "TRUE", !is.na(strain)) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))
```


## Part 4: using our controls to train clustering of our 24hr assay

### Getting classes from our own (green) data

```{r}
raw_24hr_dauers <- easysorter::read_data("data/raw/20220505_dauerProtocol4") # FOR NOW USE THIS ASSAY FOR SETTINGS

proc_dauers2 <- as.data.frame(raw_24hr_dauers[1]) %>%
  dplyr::mutate(well = paste(plate,row,col, sep = ""))

# EM clustering in DI condition
df_24hr_raw <- proc_dauers2 %>%
  dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(strain %in% c("JU346", "ECA1089", "ECA1119", "ECA1121", "NIC166", "ECA1091", "ECA1120", "ECA1122")) %>%
  dplyr::filter(condition %in% c("control", "20uM_ascr5", "10uM_ascr5"))
  #dplyr::filter(TOF>100)

df_24hr_green <- df_24hr_raw[c("TOF", "green", "EXT")]
ret.em_K4_green_24hr <- EMCluster::init.EM(df_24hr_green, nclass = 4, method = "em.EM")
df_24hr_green_merge_K4 <- data.frame(df_24hr_raw, ret.em_K4_green_24hr[6])

# plot to define classes
ggplot(df_24hr_green_merge_K4) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3)+
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title = "24hr Dauer Assay 4 Class Training (green)") +
  theme_bw() + 
  ylim(0, 300) +
  xlim(0, 600)
```

```{r}
# Caution!!! factor label changes everytime!!!
df_24hr_green_merge_K4$class <- factor(df_24hr_green_merge_K4$class, levels=c(1,4,2,3,5,6), labels=c("dauer","non-dauer", "other1", "other2"))

df_24hr_green_merge_K4$class <- fct_collapse(standrd_green_merge_K4$class, dauer = c("dauer"), nondauer = c("non-dauer1"), other = c("other1","other2"))

# replot with new factor names
df_24hr_plot_green <- df_24hr_green_merge_K4 %>%
  #dplyr::filter(!is.na(strain)) %>%
  #dplyr::filter(class %in% c("dauer", "non-dauer")) %>%
  ggplot(.) +
  aes(x = TOF, y = green, scale = T, color = as.factor(class)) +
  geom_point(size=1, alpha = 0.3) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(~condition) +
  labs(title = "24hr Dauer Assay 4 Classes (green)", color = "") +
  theme_bw() +
 #ylim(0, 300) +
  xlim(0, 600)

df_24hr_plot_green

cowplot::ggsave2(ours_plot_green, filename = "plots/24hr_green_classes4.png", width =5, height = 5)
```

